---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

````{r setup, include = FALSE}
print_yaml <- function(filename) {
  cat("```yaml", readLines(filename), "```", sep = "\n")
}
````


# test-app

[![RStudio community](https://img.shields.io/badge/community-github--actions-blue?style=social&logo=rstudio&logoColor=75AADB)](https://community.rstudio.com/new-topic?category=Package%20development&tags=github-actions)

This action tests a Shiny application (or Shiny-based document) using the [`{shinytest2}`](https://github.com/rstudio/shinytest2) package. The `${{ inputs.app-dir}}/tests/testthat/_snaps` directory will be uploaded if enabled (default).

# Usage

Inputs available:

- `app-dir` - default `"."`. Directory of Shiny application (or Shiny-based document) to test
- `upload-snapshots` - default `true`. Whether to upload all testthat snapshots as an artifact.

Typical (single app testing) GHA step usage:

``` yaml
- uses: rstudio/shinytest2/actions/test-app@actions/v1
  with:
    app-dir: |
      dir/to/app
```

Multiple Apps can be tested by supplying multiple directories to `app-dir` separated by a newline:

```yaml
- uses: rstudio/shinytest2/actions/test-app@actions/v1
  with:
    app-dir: |
      dir/to/app1
      dir/to/app2
```


# Example workflows

These workflows are a good building block / starting point for testing your Shiny applications. You may need to alter what is actually executed to fit your needs.

- [`test-app-description`](#dependencies-in-description-file) - An example CI workflow to test your Shiny application given you are using a `DESCRIPTION` file to state your dependencies.
- [`test-app-renv`](#dependencies-managed-by-renv) - An example CI workflow to test your Shiny application given you are using `{renv}` to manage your dependencies.
- [`test-app-package`](#app-within-package-structure) - An example CI workflow to test a Shiny application within your local R package. It is recommended to use `app-dir` input to set the location of your Shiny application.


## Dependencies in DESCRIPTION file

Install workflow using:
```r
usethis::use_github_action(
  url = "https://github.com/rstudio/shinytest2/raw/main/actions/test-app/example-test-app-descrption.yaml",
  save_as = "test-app-description.yaml"
)
```

Workflow contents:

```{r echo = FALSE, results = "asis"}
print_yaml("example-test-app-description.yaml")
```

## Dependencies managed by `{renv}`

Install workflow using:
```r
usethis::use_github_action(
  url = "https://github.com/rstudio/shinytest2/raw/main/actions/test-app/example-test-app-renv.yaml",
  save_as = "test-app-renv.yaml"
)
```

Workflow contents:
```{r echo = FALSE, results = "asis"}
print_yaml("example-test-app-renv.yaml")
```

## App within Package structure

If you are using a local package and want to test an application within the same repo, this may be a good option. However, it is recommended to try to execute these tests as a part of your standard `{testthat}` testting.

Install workflow using:
```r
usethis::use_github_action(
  url = "https://github.com/rstudio/shinytest2/raw/main/actions/test-app/example-test-app-package.yaml",
  save_as = "test-app-package.yaml"
)
```

Workflow contents:
```{r echo = FALSE, results = "asis"}
print_yaml("example-test-app-package.yaml")
```

# Development

The `test-app` action uses a sliding git tag that follows the pattern `actions/vX`, e.g. `actions/v1`. For historical reasons, we also support the `v1` tag, but future versions will only be available under the `actions/vX` tag.

The `test-app` action release cycle is not dependent on the shinytest2 package cycle. When changes are made to the `test-app` action, you should force-update the current sliding tag version:

```bash
git tag -f v1            # update historical v1 tag
git tag -f actions/v1    # update sliding tag
git push --tags --force  # push tag to github
```

# License

The scripts and documentation in this project are released under the [MIT License](LICENSE)

# Contributions

Contributions are welcome!
