---
title: "Monkey-testing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{monkey-testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(shinytest2)
library(shiny)
```

# Monkey (headless) testing with `{shinytest2}`

Most people will use `{shinytest2}` with the plug and play `record_test()` app, which is very convenient if you are
not familiar with JavaScript. Under the hood, `record_test()` generates a R script composed of a series of 
instructions that manipulates the app to automate testing on CI/CD environments. 

Monkey testing is widely used by web developers to check the application robustness, particularly in apps with large
number of inputs. The goal is ultimately to try to break the app by triggering unexpected combinations. Most available libraries
are JS-based such as [gremlins.js](https://github.com/marmelab/gremlins.js), traditionally combined with JS-based global testing libraries like [Puppeteer](https://pptr.dev/), but can work with `{shinytest2}` as well. 

In this vignette, we'll provide a more thorough overview of the `AppDriver` R6 class (like for [`{shinytest}`](https://rstudio.github.io/shinytest/articles/in-depth.html)), which allows the developer
to programmatically control the app. We'll see how we can seemlessly benefit from gremlins.js with only few lines of code.

## Initialize the driver

We consider a simple app composed of a slider and a plot output:

```r
ui <- fluidPage(
  sliderInput("obs", "Number of observations:",
    min = 0, max = 1000, value = 500
  ),
  plotOutput("distPlot")
)

# Server logic
server <- function(input, output) {
  output$distPlot <- renderPlot({
    hist(rnorm(input$obs))
  })
}

# Complete app with UI and server components
shinyApp(ui, server)
```

The driver may be initialized with:

```r
headless_app <- AppDriver$new(
  #load_timeout = 10,
  path = "<PATH_TO_APP>",
  name = "monkey-test",
  shiny_args = list(port = 3515)
)
```

Note the `shiny_args` slot allowing you to pass custom options to `shiny::runApp()` such as the port, which
might be useful if your organization restricts port number. `load_timeout` defaults to 10s and 20s locally and during CI/CD, respectively. 
Therefore, if your app takes longer to launch, you can change this value. Keep in mind that an app taking more than 20s to launch is generally
under-optimized and would require specific care such as profiling and refactoring. 

AppDriver starts a Chrome-based headless browser. Interestingly, if you need specific [flags](https://peter.sh/experiments/chromium-command-line-switches/) that are not
available by default in `{shinytest2}`, you can pass them before instantiating the driver:

```r
chromote::set_chrome_args(
  c(
    chromote::default_chrome_args(),
    # Custom flags: see https://peter.sh/experiments/chromium-command-line-switches/
)
```

Some flags are considered by default, particularly `--no-sandbox` which is apply only on CI/CD (Chrome won't start
without) such that you don't need to worry too much about this.

## Injecting gremlins.js

The next steps consists of injecting the gremlins.js dependency in the DOM so that we can unleash the horde.

### Easy way
The easiest way is to call:

```r
plop$execute_js("
  (function() {
    var s = document.createElement('script');
    s.src = 'https://unpkg.com/gremlins.js';
    document.body.appendChild(s);
  })();"
)
```

This creates a `script` tag pointing to the right CDN (content delivery network: optimized server to store libraries) and inserts
it at the end of the body.

To test whether everything worked well, we can dump the DOM, look for the scripts. We should be able to find
gremlins.js and try to call `typeof window.gremlins`, which should return an object:

```r
plop$get_html("script", outer_html = TRUE)
plop$execute_js("return typeof window.gremlins")
```

If `undefined` is returned instead, it means something went wrong. This is generally because the JS code
is clocked by the network. In this case, let's see what we can do in the next part.

### Local way
