---
title: "Using shinytest2 with R packages"
output: rmarkdown::html_vignette
editor_options:
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Using shinytest2 with R packages}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
library(shiny)
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

For R packages that have Shiny applications, there are generally two ways that the applications will be present in the package. The first is to have an `app.R` in a subdirectory of `inst/`. The second way is to have a function which returns a Shiny app object.

<!-- See https://github.com/rstudio/shinytestPackageExample for an example package. This document closely mirrors the content of that package: it explains how to set up tests for both types of applications, and then has a section that describes some setup tasks that are common to both types. -->

> 🚧 Using `{chromote}` in CRAN tests
>
> `{chromote}` utilizes the Google Chrome application installed on testing
> machines, which can change over time. CRAN has requested that no package utilize
> `{chromote}`'s functionality when testing on CRAN. This is because the package
> may not work as expected if the version of Google Chrome changes.
>
> To address this, `{shinytest2}` will call `testthat::skip_on_cran()` when creating an `AppDriver` during CRAN testing.
>
> To learn more about CRAN's reasoning and how to test your package using CI,
> please visit `{chromote}`'s [Using `{chromote}` in CRAN tests](https://rstudio.github.io/chromote/articles/example-cran-tests.html).



## Applications in `inst/`

An application could live in a subdirectory of `inst/`, as shown below:

```
/
├── DESCRIPTION
├── NAMESPACE
├── R
├── inst
│   └── sample_app
│       ├── app.R
│       └── tests
│           ├── testthat
│           │   ├── _snaps
│           │   │   └── shinytest2
│           │   │       └── 001.json
│           │   └── test-shinytest2.R
│           └── testthat.R
└── tests
    ├── testthat
    │   └── test-inst-apps.R
    └── testthat.R
```

In this case, you can run `record_test()` and `test_app()` as normal. After you create and run the tests, there will be a `tests/` subdirectory in the application directory that stores the test scripts and results.

Since we are using `{testthat}` for automated tests, you would create a test driver script in `tests/testthat/`. In this example, it's named `test-inst-apps.R` and contains the following:

```{r}
# File: tests/testthat/test-inst-apps.R
library(shinytest2)

test_that("sample_app works", {
  # Don't run these tests on the CRAN build servers
  skip_on_cran()

  appdir <- system.file(package = "exPackage", "sample_app")
  test_app(appdir)
})
```

If the application directory is not meant to be public, it can also be located in `./tests/testthat/apps`. `{shinytest2}` does this with many application and has `appdir` above point to the relative path to the application.

<!-- For more information about this script, see the [Test driver script notes](#Test-driver-script-notes) section below. -->


## Application objects created by functions

The second way have an application in an R package is by having a function that returns a Shiny application object. In this example, there's a function `hello_world_app()`, which lives in `R/hello-world.R`:

```
/
├── .Rbuildignore
├── DESCRIPTION
├── NAMESPACE
├── R
│   └── hello-world.R
└── tests
    ├── testthat
    │   ├── _snaps
    │   │   └── app-function
    │   │       └── 001.json
    │   └── test-app-function.R
    └── testthat.R
```

The function simply returns an object from `shinyApp()`:

```{r}
# File: R/hello-world.R

hello_world_app <- function() {
  utils::data(cars)
  shinyApp(
    ui = fluidPage(
      sliderInput("n", "n", 1, nrow(cars), 10),
      plotOutput("plot")
    ),
    server = function(input, output) {
      output$plot <- renderPlot({
        plot(
          head(cars, input$n),
          xlim = range(cars[[1]]),
          ylim = range(cars[[2]])
        )
      })
    }
  )
}
```

Once we have the object, it can be supplied directly to `AppDriver$new()`.

```{r}
# File: tests/testthat/test-app-function.R

test_that("hello-world app initial values are consistent", {
  # Don't run these tests on the CRAN build servers
  skip_on_cran()

  shiny_app <- hello_world_app()
  app <- AppDriver$new(shiny_app, name = "hello")

  app$expect_values()
})
```

To help create tests, you can call `record_test()` on your shiny application object directly. Unfortunately, the test file will not be able to be saved. Instead, the test commands can be copied into a test script manually.


<!-- For more information about this script, see the [Test driver script notes](#Test-driver-script-notes) section below. -->


## Running Applications with a function

Some packages, such as `{golem}`, provide a function that runs or returns a shiny application.  In this case, you can create a wrapper function to perform two actions: load the package and return an app. For example, if your package is named `mypkg` and the function that runs the app is `run_test_app()`, you can create a wrapper function like this:

```{r}
run_mypkg_app <- function() {
  library(mypkg)
  run_test_app()
}

app <- AppDriver$new(run_mypkg_app, name = "mypkg-app")
```

For shorthand, you can also use the package function directly:

```{r}
app <- AppDriver$new(run_test_app, name = "mypkg-app")
```

When using a function to initalize the `AppDriver`, `{shinytest2}` will automatically call `pkgload::load_all()` when you execute `library(<pkg>)` or `require(<pkg>)` to load your dev package, `<pkg>`, in the background R process. Similar to `{mirai}`, there is no carryover to the execution within the background R process. All packages may need to be `library()`'ed again.

`{golem}` provides a template `run_app()` function. `{golem}` app authors can use the shorthand:
```{r}
app <- AppDriver$new(run_app)
```


## Other setup steps

There are a few steps that are needed for both types of tests.

It is recommended to call `shinytest2::use_shinytest2()` to enable different test config set-ups.

You will need to add `{shinytest2}` to the `Suggests` section in your `DESCRIPTION` file.

```
Suggests:
    shinytest2
```

When all of these items are in place, you can test your package using `testthat::test_local()` or by running ``R CMD check`` on your package. If you are using the RStudio IDE, you can also run Build -> Test Package or Build -> Check Package.

`{shinytest2}` **no longer** requires that your package to be *installed* when testing. It will load from your local path. [`testthat::test_local()`](https://testthat.r-lib.org/reference/test_package.html) (and related wrappers) eventually call `pkgload::load_all()` to temporarily source the local R package in the foreground for unit testing. `{shinytest2}` has made the stance that App authors **must** call `library(<pkg>)` or `require(<pkg>)` on their package within their App. This ensures that the package is loaded in the background R process that runs the Shiny application. When `library*()` is called, `{shinytest2}` will automatically execute `pkgload::load_all()` to load the package's source code.

## How should I test multiple applications?

You can call `shinytest2::test_app()` multiple times within a test script. It does not need to be wrapped within a `testthat::test_that()` call.

`{shinytest2}` tests many internal apps using the code similar to the code below:

``` r
# ./tests/testthat/test-apps.R

app_dirs <- fs::dir_ls(testthat::test_path("apps"))
lapply(app_dirs, function(app_dir) {
  shinytest2::test_app(app_dir)
})
```


## Continuous integration

If you would like your package to be tested with every commit, you can set it up with GitHub Actions. Please see [Using shinytest2 with continuous integration](./use-ci.html) for inspiration.
